---
title: "p8105_hw6_yk2959"
author: "Kiran Kui"
date: "2022-11-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(p8105.datasets)
library(viridis)
library(readxl)
library(dplyr)
library(purrr)
library(ggplot2)
```


# Problem 2

```{r}
homicide_df = read_csv("./data/homicide-data.csv") %>%
  janitor::clean_names() 

skimr::skim(homicide_df)
```

I have imported the homicide dataset from the github repository and tidied the names of the variables in the dataset. 

## Brief introduction of the raw homicide dataset 

The raw dataset contains `r nrow(homicide_df)` rows and `r ncol(homicide_df)` columns, with each row representing a single record of homicide in 50 large U.S. cities. 

Variables include, date of homicide, victim first and last name, victims' race, age, sex and the city/state that the homicide was located, as well as the lattitude and longitude of the homicide location, given in `r colnames(homicide_df)`. 

```{r}
homicide_df_clean_1 = homicide_df %>% 
  unite(city_state, c(city, state), sep = ", ", remove = FALSE) %>% 
  select(-city, -state) %>% 
  mutate(
  resolved = ifelse(
  disposition %in% c('Closed without arrest', 'Open/No arrest'), 0, 1)) %>% 
    filter(!city_state %in% c('Dallas, TX', 'Phoenix, AZ', 'Kansas City, MO', 'Tulsa, AL')) %>%    filter(victim_race %in% c('Black', 'White')) %>%
  mutate(victim_age = as.numeric(victim_age))

homicide_df_clean_1
```

I create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. I then omitted cities Dallas, TX; Phoenix, AZ; and Kansas City, MO because these don’t report victim race and also omit Tulsa, AL because this is a data entry mistake. I limited my analysis those for whom victim_race is white or black.I converted victim_age from a character to a numeric variable. 

```{r}
fit_logistic = homicide_df_clean_1 %>% 
  filter(city_state %in% c('Baltimore, MD')) %>% 
  glm(resolved ~ victim_age + victim_sex + victim_race, data = ., family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    odds_ratio = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, odds_ratio, lower_CI, upper_CI) %>%
  filter(term == "victim_sexMale") %>% 
  knitr::kable(digits = 3)

fit_logistic

```

For the city of Baltimore, MD, I used the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 

I saved the output of glm as an R object and applied the broom::tidy to this object then obtained the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}

homicide_df_clean_2 = 
  homicide_df_clean_1 %>% 
  nest(data = -city_state) %>% 
mutate(
  all_countries = 
    map(.x = data, ~glm(glm(resolved ~ victim_age + victim_race + victim_sex, data = .x, family = binomial()))),
  result = map(all_countries, broom::tidy) 
  ) %>% 
select(city_state, result) %>% 
 unnest(result) %>% 
  mutate(
    odds_ratio = exp(estimate),
    lower_CI = exp(estimate - 1.96 * std.error),
    upper_CI = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, odds_ratio,lower_CI, upper_CI) %>% 
  filter(term == "victim_sexMale") %>% 
  mutate(city_state = fct_reorder(city_state, odds_ratio)) 

ggplot_q2 = 
  homicide_df_clean_2 %>% 
  ggplot(aes(x = city_state, y = odds_ratio)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  labs(title = "Plot showing the estimated ORs and CIs for each city",
    x = "City, State",
    y = "Odds Ratio")

ggplot_q2

```

I ran a glm for each of the cities in your dataset, and extracted the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

I created a plot that shows the estimated ORs and CIs for each city and organized cities according to estimated OR, and comment on the plot.



